Project assignment - Tahoe-LAFS with SHA-3 candidates

### 2010-11-10 paper/main.tex

\section{Cryptographic Hash Functions}
% Get a source for this %
A Cryptographic Hash function is a deterministic mathematical procedure which
takes an arbritary block of data and outputs a fixed-size bit string. The
output is refered to as the hash value, message digest or simply digest.
Another property of a cryptographic hash function is that a small change in the
input data (just 1 bit) should completly change the output of the hash
function. In other words it should be infeasible to find the reverse of a
cryptographic hash function. It should also be infeasible to find two blocks of
data which produce the same hash value (a collision).

\subsection{NIST SHA-3 Competition}
The \ac{SHA} version 3 is a comming standard set to superceed the current
standards \ac{SHA}-1 and the \ac{SHA}-2 family. The hash function to become
\ac{SHA}-3 will be decided by \ac{NIST} and choosen between the submitted
contestants to the \ac{NIST} hash competition. The current status of the 
competition is officially called Round 2, with 14 of 64 candidates having 
survived round 1\cite{s_fedreg}.

\subsubsection{\ac{NIST} evaluation criterias for \ac{SHA}-3}

\paragraph{Security} is the most important criteria for the SHA-3
candidates\cite{s_nistround2}. The full description of security criterias are
mentioned in \cite{s_fedreg}, however the most noteworthy criteria in regards
to Tahoe-\ac{LAFS} is that \ac{SHA}-3 candidates are required to have
resistance against length-extension attacks. Both \ac{SHA}-1 and the \ac{SHA}-2
family are vulnerable to this attack\cite{s_mdre}.

\paragraph{Cost and Performance} are considered to be the 2nd most important
criteria. The absolute minimum for performance is that the SHA-3 candidate
should be faster than the functions in the \ac{SHA}-2 family. Cost is a measure
of how much memory an implementation requires in software, both the 
implementation itself and the use of \ac{RAM} during runtime. Another measure
of cost is how many logic gates it takes to implement the function in hardware.

\paragraph{Algorithm and implementation characteristics} is the third criteria.
By this criteria \ac{NIST} emphasis that algorithms with greater flexbility
will be given preference over other algorithms\cite{s_nistround2}. By flexible
they imply the possiblity of the function to run efficiently on a varity of
plattforms, to use parallelism or instruction set extensions. Another key point
of a flexible hash function is that it should have a simple and elegant design
to encourage understanding, analysis and design confidence.

\subsection{\ac{SUPERCOP}}
% TODO: Mention that some SUPERCOP candidates are missing UPDATE-functionality.
\ac{SUPERCOP} is a toolkit developed by VAMPIRE lab for measuring performance
of cryptographic software\cite{s_supercop}. In relation to hash functions
\ac{SUPERCOP} measures the following:

\begin{itemize}
    \item Time to hash a very short packet of data
    \item Time to hash a typical-size Internet packet
    \item Time to hash a long message
    \item Length of the hash output
\end{itemize}

The round 2 \ac{SHA}-3 candidates are all included in the toolkit, with a
number of different optimalizations for each candidate. Optimalizations range
from 32-/64-bit optimalizations, the use of extended instruction sets such as
\ac{SSE2} and \ac{SSSE3}, optimalizations for different number of cores and
others. Also the toolkit will try different compiler-optimalizations to get the
best results possible for each function.
% Skal jeg sitere kildekode her?
% TODO: Insert image to visualize the forest of different implementations.

The benchmarking results for the \ac{SUPERCOP} toolkit of \ac{SHA}-3 candidates
and \ac{SHA}-2 functions on a number of different plattforms and architectures 
are available at
\footnote{\url{http://bench.cr.yp.to/results-sha3.html}}

>>>>

\section{Cryptographic Hash Functions}

A cryptographic hash function is a deterministic mathematical procedure which
takes an arbitrary block of data and outputs a fixed-size bit string. The output
is referred to as the hash value, message digest or simply digest.
Another property of a cryptographic hash function is that the smallest change in
the input data (e.g. one bit) should completely change the output of the hash
function. In other words it should be infeasible to find the reverse of a
cryptographic hash function \cite{stallings}. It should also be infeasible to
find two blocks of data which produce the same hash value (a \emph{collision}).

\subsection{NIST SHA-3 Competition}
The \ac{SHA} version 3 is a coming standard set to supersede the current
standards that the \ac{SHA}-1 and the \ac{SHA}-2 family has become. The hash
function that will be known as \ac{SHA}-3 will be decided by the \ac{NIST} and
chosen between the submitted contestants to the \ac{NIST} hash competition.
At the time of writing, the current status of the competition is officially
called Round 2, with 14 of 64 candidates having ``survived'' Round 1
\cite{s_fedreg}.

\subsubsection{\ac{NIST} evaluation criteria for \ac{SHA}-3}

\paragraph{Security.} The most important criteria for the SHA-3
candidates\cite{s_nistround2} is security. The full description of security
criteria are mentioned in \cite{s_fedreg}, however the most noteworthy criterion
in regards to Tahoe-\ac{LAFS} is that \ac{SHA}-3 candidates are required to have
resistance against length-extension attacks. Both \ac{SHA}-1 and the \ac{SHA}-2
family are vulnerable to this attack\cite{s_mdre}, and this requires Tahoe to
run the algorithm twice, as described by .
%TODO: cite Practical cryptography/Niels Ferguson, Bruce Schneier

\paragraph{Cost and Performance.} Cost and Performance are considered to be the
2nd most important criterion. The absolute minimum for performance is that the
SHA-3 candidate should be faster than the functions in the \ac{SHA}-2 family.
Cost is a measure of how much memory an implementation requires in software,
both the implementation itself and the use of \ac{RAM} during runtime. Another
measure of cost is how many logic gates it takes to implement the function in
hardware.

\paragraph{Algorithm and implementation characteristics.} By this 3rd criteria,
\ac{NIST} emphasis that algorithms with greater flexibility will be given
preference over other algorithms \cite{s_nistround2}. By flexible they imply the
possibility of the function to run efficiently on a variety of platforms, and to
use parallelism or instruction set extensions. Another key point of a flexible
hash function is that it should have a simple and elegant design to encourage
understanding, analysis and design confidence.

\subsection{\ac{SUPERCOP}}
% TODO: Mention that some SUPERCOP candidates are missing UPDATE-functionality.
\ac{SUPERCOP} is a toolkit developed by VAMPIRE lab for measuring performance
of cryptographic software \cite{s_supercop}. In relation to hash functions
\ac{SUPERCOP} measures the following:

\begin{itemize}
    \item Time to hash a very short packet of data.
    \item Time to hash a typical-size Internet packet.
    \item Time to hash a long message.
    \item Length of the hash output.
\end{itemize}

The Round 2 \ac{SHA}-3 candidates are all included in the toolkit, with a number
of different optimizations for each candidate. Optimizations range from
32-/64-bit specific implementations, the use of extended instruction sets such
as \ac{SSE2} and \ac{SSSE3}, optimizations for different number of cores and
others. The toolkit will also try different compiler optimizations to get the
best results possible for each function.
% TODO: Insert image to visualize the forest of different implementations.

The benchmarking results for the \ac{SUPERCOP} toolkit of the \ac{SHA}-3
candidates and \ac{SHA}-2 functions on a number of different plattforms and
architectures are available at
\footnote{\url{http://bench.cr.yp.to/results-sha3.html}}
% TODO: Link to appendix with tahoe04 SUPERCOP results.
